Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis

### 2025-08-30 Paper/introduction.tex
More on intro

@@

>>>>

Owing to the security-critical nature of many applications of string analysis
(e.g. analysis of web applications against XSS vulnerabilities), there is an 
increasing need of trustworthy implementations in such application domains. 
Unfortunately, string analysis implementations are often quite buggy. For
example, even state-of-the-art SMT solvers over the theory of strings --- 
which were used at AWS, esp. for analyzing Role-Based Access Control (RBAC) 
policies --- were found to be buggy through such techniques as fuzzing
\cite{DBLP:conf/cav/BlotskyMBZKG18,BM20,Mansur20}. This has motivated
initial formalization efforts for the theory of strings in Isabelle/HOL, e.g., a verified solver
\cite{cpp/KanLRS22}
for handling a small fragment of SMT-LIB 2.6 unicode theory of strings
consisting only of equality, concatenation and regular constraints, and a
formalization of the semantics of this theory \cite{verified-verifying}.

Despite these, existing verified implementations for string analysis do not
handle the following indispensable string operations for analysis of web
applications (e.g. see
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}):
\emph{string transductions} and \emph{replaceall} operators. In particular, 
string transductions are used frequently in web applications (e.g. for
sanitization, as well as implicitly applied transductions, for instance
through \texttt{innerHTML}). In particular, SMT-LIB string constraints that 
arise from such applications require the \emph{replaceall} operator (part of
SMT-LIB), which is currently not handled by any existing verified
implementation. This paper tackles the formalization of SFA/SFT and
their associated algorithms, as well as applications to analysis of sanitizers
and verified solver for string constraints. \textcolor{red}{Updated up to here
so far}

\OMIT{
SFA and SFT (e.g. 
analysis of web applications, e.g., against XSS vulnerabilities),

@@

}
